Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Clique Neighbourhood Constraints

### 2021-01-18 paper/paper.tex
Sonja's comments from Overleaf

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
\usepackage{ijcai21}
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}

\usepackage{cleveref}
\crefname{figure}{figure}{figures}
\Crefname{figure}{Figure}{Figures}

@@

A \emph{graph homomorphism} is a function between two graphs that maps adjacent vertices to adjacent
vertices. An injective graph homomorphism is more commonly known as a \emph{subgraph isomorphism},
whilst a \emph{locally injective homomorphism} is one which is injective when restricted to any
individual vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}. Finding any of these three kinds of homomorphism between
two given graphs is \NP-complete.  The subgraph isomorphism problem arises in many practical
applications, including ??, and as such there has been a lot of research into finding algorithms
which perform well empirically. Most approaches are based either upon very fast but simple
backtracking algorithms,
\cite{DBLP:journals/pami/CordellaFSV04,DBLP:journals/bmcbi/BonniciGPSF13,DBLP:conf/gbrpr/CarlettiFSV17}
which often but not always perform well on very easy instances, or upon constraint programming
algorithms
\cite{DBLP:journals/constraints/ZampelliDS10,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/ArchibaldDHMP019},
which have higher startup costs but that perform vastly better on harder instances and much more
consistently on easy instances \cite{DBLP:journals/jair/McCreeshPST18,DBLP:conf/gbrpr/Solnon19}. The
current state of the art is the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}, which is a
dedicated constraint programming solver for subgraph-finding problems. Much of its performance comes
from inference strategies based upon degrees and neighbourhood degree sequences
\cite{DBLP:journals/constraints/ZampelliDS10}, counting paths between vertices
\cite{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, and cardinality reasoning
\cite{DBLP:journals/ai/Solnon10}, which can be used to eliminate many infeasible candidate
assignments without search.

One might hope that this inference would also be helpful for other graph homomorphism problem
variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
the locally injective case, and that simple distance filtering is valid for all homomorphisms.
However, we also show that \emph{none} of the other strategies are valid for finding homomorphisms
where there is no injectivity requirement. Finally, we introduce a new filtering technique that is
based upon finding a maximum clique in the neighbourhood of each pattern vertex, and prove that this
technique is valid even in the general case. Although this new filtering technique involves solving
many additional \NP-complete problems as a preprocessing step, we demonstrate that it is effective
in practice, particularly for the non-injective problem where filtering allows a constraint
programming algorithm to solve instances over eight hundred times faster in aggregate, over a
collection of over fourteen thousand standard benchmark instances. This shows, for the first time,
that modern constraint programming techniques can be practical for less constrained graph
homomorphism finding problems; previous algorithmic approaches have focused instead upon worst-case
computational bounds
\cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
especially important because many current applications use subgraph isomorphism solvers only because
they perform well using off the shelf solvers, rather than because they exactly match ideal domain
requirements ??cite.

@@

It is easy to verify that the first of these two properties is valid for \emph{any} homomorphism.
The second property does not hold for homomorphisms in general: for example, a pair of vertices
connected by three paths of length two may be mapped onto a pair of vertices connected by a single
path of length two (see \cref{figure:counterexamples}, top right). Local injectivity is trickier.
The second property does not hold in full generality, because, for example, a path of length three
may be mapped onto a triangle whilst respecting local injectivity (\cref{figure:counterexamples},
bottom left).  However, we observe that a weaker result holds:

\begin{proposition}[paths of length two are preserved by locally injective
    homomorphisms]\label{proposition:paths}For the problem of finding a locally injective graph
    homomorphism $i$ from a graph $G$ to a graph $H$, for any pair of vertices $v, w \in
    \vertexset(G)$, if there are at least $k$ simple paths of length exactly two between $v$ and
    $w$, then there must be at least $k$ simple paths of length exactly two between $i(v)$ and
    $i(w)$.
\end{proposition}

@@

\begin{corollary}\Cref{proposition:clique} holds for locally injective graph homomorphisms and for
    subgraph isomorphisms even if loops are present in the second graph.
\end{corollary}

@@

We calculate the maximum clique size for the neighbourhood of a first pattern vertex $p_1$.  Having
found a maximum clique $\omega_1$ for a vertex $p$, for every other vertex $q \in \omega_1$, we
remember that the maximum clique for $q$'s neighbourhood has at least $|\omega_1|$ vertices. We then
proceed to find a maximum clique $\omega_2$ for a second pattern vertex $p_2$; however, if $p_2$ was
a member of $\omega_1$, we initialise the clique solver with a lower bound of $n$ rather than
starting from scratch. We then find the maximum clique size for a third pattern vertex $p_3$, using
whichever lower bound of $|\omega_1|$ and $|\omega_2|$ is better (if both are valid), and so on.

We then move on to the target vertices, using a similar caching routine. Rather than calculating a
clique size for every single target vertex, we only calculate a value for target vertices which are
present in at least one variable's domain, after other unary constraints have been applied.
(Depending upon the problem variant, implied constraints like degree and neighbourhood degree
sequence can eliminate some target vertices from the domains of all pattern variables.)
Additionally, we do not require the maximum clique solver to run to completion and guarantee that it
has found a maximum clique. Instead, we allow it to stop as soon as it has found a clique with as
many vertices as the largest pattern clique. This is useful in practice because it may be very hard
to decide whether a particular target vertex has neighbourhood clique size of, say, 15 or 16, but if
the largest pattern vertex has a neighbourhood clique size of only 5 then this is irrelevant.

>>>>

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
\usepackage{ijcai21}
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}
\usepackage[textsize=tiny,textwidth=2cm,shadow]{todonotes}

\newcommand{\skcom}[1]{\todo[color=orange!50!white]{{S}: #1}}
\newcommand{\rcom}[1]{\todo[color=gray!50!white]{{R}: #1}}%reviewers said
\usepackage{cleveref}
\crefname{figure}{figure}{figures}
\Crefname{figure}{Figure}{Figures}

@@

A \emph{graph homomorphism} is a function between two graphs that maps adjacent vertices to adjacent
vertices. An injective graph homomorphism is more commonly known as a \emph{subgraph isomorphism},
whilst a \emph{locally injective homomorphism} is one which is injective when restricted to any
individual vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}.\skcom{It seems a little bit strange to start with a definition. How about we describe it more wordy in the introduction? Or maybe I would start motivating with subgraph isomorphism(which has an easy definition i words) and the fact that this doesnt ideally match many causes... and then go to the definition of general homomorphism? } Finding any of these three kinds of homomorphism between
two given graphs is \NP-complete\skcom{cite a proof of this}.  The subgraph isomorphism problem arises in many practical
applications, including ??, and as such there has been a lot of research into finding algorithms
which perform well empirically. Most approaches are based either upon very fast but simple
backtracking algorithms,
\cite{DBLP:journals/pami/CordellaFSV04,DBLP:journals/bmcbi/BonniciGPSF13,DBLP:conf/gbrpr/CarlettiFSV17}
which often but not always perform well on very easy instances, or upon constraint programming
algorithms
\cite{DBLP:journals/constraints/ZampelliDS10,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/ArchibaldDHMP019},
which have higher startup costs but that perform vastly better on harder instances and much more
consistently on easy instances \cite{DBLP:journals/jair/McCreeshPST18,DBLP:conf/gbrpr/Solnon19}. The
current state of the art is the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}, which is a
dedicated constraint programming solver for subgraph-finding problems. Much of its performance comes
from inference strategies based upon degrees and neighbourhood degree sequences
\cite{DBLP:journals/constraints/ZampelliDS10}, counting paths between vertices
\cite{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, and cardinality reasoning
\cite{DBLP:journals/ai/Solnon10}, which can be used to eliminate many infeasible candidate
assignments without search.

One might hope that this inference would also be helpful for other graph homomorphism problem
variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
the locally injective case, and that simple distance filtering is valid for all homomorphisms.
However, we also show that \emph{none} of the other strategies are valid for finding homomorphisms
where there is no injectivity requirement. Finally, we introduce a new filtering technique that is
based upon finding a maximum clique in the neighbourhood of each pattern vertex, a
technique, that is valid even in the general case.\skcom{I wouldnt emphasize so much that "we prove" - its a common result and also pretty immediate..} Although this new filtering technique involves solving
many additional \NP-complete problems as a preprocessing step, we demonstrate that it is effective
in practice, particularly for the non-injective problem where filtering allows a constraint
programming algorithm to solve instances over eight hundred times faster in aggregate, over a
collection of over fourteen thousand standard benchmark instances. This shows, for the first time,
that modern constraint programming techniques can be practical for less constrained graph
homomorphism finding problems; previous algorithmic approaches have focused instead upon worst-case
computational bounds
\cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
especially important because many current applications use subgraph isomorphism solvers only because
they perform well using off the shelf solvers, rather than because they exactly match ideal domain
requirements ??cite.

@@

It is easy to verify that the first of these two properties is valid for \emph{any} homomorphism.
The second property does not hold for homomorphisms in general: for example, a pair of vertices
connected by three paths of length two may be mapped onto a pair of vertices connected by a single
path of length two (see \cref{figure:counterexamples}, top right). While standard injectivity prevents any kind of merging, local injectivity prevents two vertices from being merged, only if they are both in the neighbourhood of the same vertex.\skcom{added this}
So the second property does not hold in full generality, because, for example, a path of length three
may be mapped onto a triangle whilst respecting local injectivity (\cref{figure:counterexamples},
bottom left).  However, we observe that a weaker result holds:

\begin{proposition}[paths of length two are preserved by locally injective
    homomorphisms]\label{proposition:paths}For the problem of finding a locally injective graph
    homomorphism $i$ from a graph $G$ to a graph $H$, for any pair of vertices $v, w \in
    \vertexset(G)$, if there are at least $k$ simple paths of length exactly two between $v$ and
    $w$, then there must be at least $k$ simple paths of length exactly two between $i(v)$ and
    $i(w)$.
\end{proposition}\skcom{I am unsure whether not to write a general statement combining this and proposition 2.2.. but maybe that}

@@

\begin{corollary}\Cref{proposition:clique} holds for locally injective graph homomorphisms and for
    subgraph isomorphisms even if loops are present in the second graph.
\end{corollary}\skcom{Do you think we should say it or just state in the definition sections that both of these '
\textit{are} homomorphisms}

@@

We calculate the maximum clique size for the neighbourhood of a first pattern vertex $p_1$.  Having
found a maximum clique $\omega_1$ for a vertex $p$, for every other vertex $q \in \omega_1$, we
remember that the maximum clique for $q$'s neighbourhood has at least $|\omega_1|$ vertices. We then
proceed to find a maximum clique $\omega_2$ for a second pattern vertex $p_2$; however, if $p_2$ was
a member of $\omega_1$, we initialise the clique solver with a lower bound of $n$\skcom{why n?} rather than
starting from scratch. We then find the maximum clique size for a third pattern vertex $p_3$, using
whichever lower bound of $|\omega_1|$ and $|\omega_2|$ is better (if both are valid), and so on.\skcom{I think after you state the first example we can just say that in general, For vertex v, we take the largest lowerbound of maxclique values in its neighbourhood (initialise to 0)?}\skcom{Maybe explicitly say in which order we take the vertices?}

We then move on to the target vertices, using a similar caching routine. Rather than calculating a
clique size for every single target vertex, we only calculate a value for target vertices which are
present in at least one variable's domain, after other unary constraints have been applied.
(Depending upon the problem variant, implied constraints like degree and neighbourhood degree
sequence can eliminate some target vertices from the domains of all pattern variables.)
Additionally, we do not require the maximum clique solver to run to completion and guarantee that it
has found a maximum clique. Instead, we allow it to stop as soon as it has found a clique with as
many vertices as the largest pattern clique. This is useful in practice because it may be very hard
to decide whether a particular target vertex has neighbourhood clique size of, say, 15 or 16, but if
the largest pattern vertex has a neighbourhood clique size of only 5 then this is irrelevant.\skcom{I would change this bit. I think it should be clear that on the second graph we \textit{dont} find a maximum clique, but what we want (for our cliques are preserved constraint) is cliques$\geq$ max size in the pattern vertex or in general of at least the sizes we find in the pattern vertex}
